---
title: "finalEDA"
author: "Team 6"
date: "2022-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing all the relevant packages and libraries 

```{r results='markup'}
#install packages
#remotes::install_github("cran/DMwR")
#install.packages("moments")

#import libraries all
library(ggplot2)
library(moments)
library(dplyr)
library(DMwR)
``` 

Importing our dataset and checking summary
```{r results='markup'}
#importing brain dataset
data <- data.frame(read.csv("dataset.csv"))


#checking the structure of our dataset 
str(data)

#since ID won't be useful during prediction so removing that 
brainData <- select(data, -c('id'))

#checking top5 entries in our dataset 
head(brainData)
```

Checking the summary of our dataset
```{r results='markup'}
summary(brainData)
```

Since most of the varibales our categorical, converting them to factors
```{r results='markup'}
#Converting gender,hypertension,heart_disease,ever_married,work_type,Residence_type,smoking_status,stroke to factors
brainData$gender = factor(brainData$gender)
brainData$hypertension = factor(brainData$hypertension)
brainData$heart_disease = factor(brainData$heart_disease)
brainData$ever_married = factor(brainData$ever_married)
brainData$work_type = factor(brainData$work_type)
brainData$Residence_type = factor(brainData$Residence_type)
brainData$smoking_status = factor(brainData$smoking_status)
brainData$stroke = factor(brainData$stroke)
str(brainData)
summary(brainData)
```


*We can see missing values in bmi, smoking_status \
**Class Imbalance problem** \
we have 42617 observations for stroke = 0 and 783 observations for stroke =1 \
\ 
Cleaning the dataset


Removing missing Values for smoking status 
```{r results='markup'}
cleanData = brainData
cleanData[cleanData == ''] <- NA
nrow(cleanData)
cleanData <- subset(cleanData, !is.na(smoking_status))
nrow(cleanData)
```
Analysis for BMI NULL values.
```{r, results='markup'}
ggplot(data=brainData, aes(bmi)) +
geom_histogram(col="black",
fill="blue") +
labs(x="bmi", y="frequency") +
labs(title="plot of bmi")
```
```{r, results='markup'}
tempdf <- subset(brainData, !is.na(bmi))  # removing NA values from bmi and storing it in tempdf
density_bmi <- density(tempdf$bmi)
plot(density_bmi)    #plotting the density curve of bmi values

ggplot(data = tempdf, mapping = aes(y = bmi)) +
  geom_boxplot() + geom_boxplot( colour="black", fill="#7777cc", outlier.colour="red", outlier.shape=8, outlier.size=2)  #plotting the boxplot for bmi values.
```
```{r, results='markup'}
skewness_value <- skewness(tempdf$bmi)
```
Skewness value: `r skewness_value`


From the density curve and skewness value, BMI is rightly skewed. So mean is not the right central tendency to fill NA values in BMI.
We will NA values in BMI using median of the data.
Note: Negative skewness value indicates left skewed and positive skewness value indicates right skewed.
The rule of thumb seems to be:

If the skewness is between -0.5 and 0.5, the data are fairly symmetrical
If the skewness is between -1 and – 0.5 or between 0.5 and 1, the data are moderately skewed
If the skewness is less than -1 or greater than 1, the data are highly skewed

Replacing Missing Values for BMI with median.
```{r results='markup'}

#removing missing values for BMI
cleanData$bmi[is.na(cleanData$bmi)] <- median(cleanData$bmi, na.rm = T) #replacing NA values with median of the remaining bmi values.
summary(cleanData)
```

Removing Outliers for BMI 
```{r results='markup'}

dim(cleanData)
quartiles_bmi <- quantile(cleanData$bmi, probs=c(.25, .75), na.rm = FALSE)
IQR_bmi <- IQR(cleanData$bmi)

 
Lower_bmi <- quartiles_bmi[1] - 1.5*IQR_bmi
Upper_bmi <- quartiles_bmi[2] + 1.5*IQR_bmi

#since all the missing values are for stroke =0 entries, we will directly drop it

cleanData <- subset(cleanData, cleanData$bmi > Lower_bmi & cleanData$bmi < Upper_bmi)

dim(cleanData)
```
```{r, results='markup'}

```


Removing Outliers for Average Glucose Level
```{r results='markup'}

dim(cleanData)
quartiles_glucose <- quantile(cleanData$avg_glucose_level, probs=c(.25, .75), na.rm = FALSE)
IQR_glucose <- IQR(cleanData$avg_glucose_level)

 
Lower_glucose <- quartiles_glucose[1] - 1.5*IQR_glucose
Upper_glucose <- quartiles_glucose[2] + 1.5*IQR_glucose

#since all the missing values are for stroke =0 entries, we will directly drop it

cleanData <- subset(cleanData, cleanData$avg_glucose_level > Lower_glucose & cleanData$avg_glucose_level < Upper_glucose)

dim(cleanData)

```

Let's check the distribution for numeric variables, bmi,age, avg_glucose_level : boxplot(outlier) and histogram(normal) => outliers **while doing avg_glucose_level and bmi => groupwise analysis as well**
3)do plots

```{r results='markup'}
```

4)show missing values for bmi and smoking_status
5)filling NA for bmi
6)removie for smoking_status 
7)remove outliers


8)smote
```{r results='markup'}

table(cleanData$stroke)
#use SMOTE to create new dataset that is more balanced
cleanData <- SMOTE(stroke ~ ., cleanData, perc.over = 2000, perc.under = 200)
#minority => 2000/100 =20 so 20*(strok1Observations=416) #upsampling 
#200/100 =2, make the majority observation 2 time the minority observations #downsampling 

#view distribution of response variable in new dataset
table(cleanData$stroke)

```
9)plot them again : qqplot => normality  

```{r results='markup'}

```

10) Comparison of age,bmi, avg_glucose_level to stroke
11) correlation plot for numeric variables

12) Plot correlation matrix for categorical variables 
13) analysis of residence_type and drop it 
14)Heart Disease: variable + age 
15)Hypertension : variable +age
16)Ever Married
17)Smoking Status
18)Gender
19)WorkType
20)All the testing anova, chi-square
21)Conclusion : major factor, what all other factors it depends on, (logistic regression test)




