---
title: "finalEDA"
author: "Team 6"
date: "2022-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing all the relevant packages and libraries 

```{r results='markup'}
#install packages
#remotes::install_github("cran/DMwR")
#install.packages("moments")

#import libraries all
library(ggplot2)
library(moments)
library(dplyr)
library(DMwR)
library(SmartEDA) 
``` 

Importing our dataset and checking summary
```{r results='markup'}
#importing brain dataset
data <- data.frame(read.csv("dataset.csv"))


#checking the structure of our dataset 
str(data)

#since ID won't be useful during prediction so removing that 
brainData <- select(data, -c('id'))

#checking top5 entries in our dataset 
head(brainData)
```

Checking the summary of our dataset
```{r results='markup'}
summary(brainData)
```

Since most of the varibales our categorical, converting them to factors
```{r results='markup'}
#Converting gender,hypertension,heart_disease,ever_married,work_type,Residence_type,smoking_status,stroke to factors
brainData$gender = factor(brainData$gender)
brainData$hypertension = factor(brainData$hypertension)
brainData$heart_disease = factor(brainData$heart_disease)
brainData$ever_married = factor(brainData$ever_married)
brainData$work_type = factor(brainData$work_type)
brainData$Residence_type = factor(brainData$Residence_type)
brainData$smoking_status = factor(brainData$smoking_status)
brainData$stroke = factor(brainData$stroke)
str(brainData)
summary(brainData)
```


*We can see missing values in bmi, smoking_status \
**Class Imbalance problem** \
we have 42617 observations for stroke = 0 and 783 observations for stroke =1 \
\ 
Cleaning the dataset


Removing missing Values for smoking status 
```{r results='markup'}
cleanData = brainData
cleanData[cleanData == ''] <- NA
nrow(cleanData)
cleanData <- subset(cleanData, !is.na(smoking_status))
nrow(cleanData)
```
Analysis for BMI NULL values.
```{r, results='markup'}
ggplot(data=brainData, aes(bmi)) +
geom_histogram(col="black",
fill="blue") +
labs(x="bmi", y="frequency") +
labs(title="plot of bmi")
```
```{r, results='markup'}
tempdf <- subset(brainData, !is.na(bmi))  # removing NA values from bmi and storing it in tempdf
density_bmi <- density(tempdf$bmi)
plot(density_bmi)    #plotting the density curve of bmi values

ggplot(data = tempdf, mapping = aes(y = bmi)) +
  geom_boxplot() + geom_boxplot( colour="black", fill="#7777cc", outlier.colour="red", outlier.shape=8, outlier.size=2)  #plotting the boxplot for bmi values.
```
```{r, results='markup'}
skewness_value <- skewness(tempdf$bmi)
```
Skewness value: `r skewness_value`


From the density curve and skewness value, BMI is rightly skewed. So mean is not the right central tendency to fill NA values in BMI.
We will NA values in BMI using median of the data.
Note: Negative skewness value indicates left skewed and positive skewness value indicates right skewed.
The rule of thumb seems to be:

If the skewness is between -0.5 and 0.5, the data are fairly symmetrical
If the skewness is between -1 and – 0.5 or between 0.5 and 1, the data are moderately skewed
If the skewness is less than -1 or greater than 1, the data are highly skewed

Replacing Missing Values for BMI with median.
```{r results='markup'}

#removing missing values for BMI
cleanData$bmi[is.na(cleanData$bmi)] <- median(cleanData$bmi, na.rm = T) #replacing NA values with median of the remaining bmi values.
summary(cleanData)
```

Removing Outliers for BMI 
```{r results='markup'}

dim(cleanData)
quartiles_bmi <- quantile(cleanData$bmi, probs=c(.25, .75), na.rm = FALSE)
IQR_bmi <- IQR(cleanData$bmi)

 
Lower_bmi <- quartiles_bmi[1] - 1.5*IQR_bmi
Upper_bmi <- quartiles_bmi[2] + 1.5*IQR_bmi

#since all the missing values are for stroke =0 entries, we will directly drop it

cleanData <- subset(cleanData, cleanData$bmi > Lower_bmi & cleanData$bmi < Upper_bmi)

dim(cleanData)
```
We won't be removing Outliers for Age and Average level because\
1) max value of age is 82, so nothing is unusual about that \
2)avg_glucose_level is losing the relationship with stroke if we remove the outliers \
When we try to upsample that difference is even more nullified \



Let's check the distribution for numeric variables, bmi,age, avg_glucose_level : boxplot(outlier) and histogram(normal) => outliers **while doing avg_glucose_level and bmi => groupwise analysis as well**

Making a separate column for average glucose level \
Here as we can see, people who have diabetes are more likely to get stroke.

```{r results='markup'}
cleanData$glucoseGroup=cut(cleanData$avg_glucose_level, c(0, 114, 140, Inf), c("Normal", "Prediabetes", "Diabetes"), include.lowest=TRUE)
#0-114 : Normal glucose level
#114-140 : Prediabetes
#140- infinity : Diabetes 

#glucoseGroup to stroke relations
#ggbarstats(data=cleanData,y=glucoseGroup,x=stroke,label="both")
#people with diabetes have a high chance of getting stroke 
df_glucose <- cleanData %>%
  group_by(glucoseGroup, stroke) %>%
  count(stroke)
#grouping data into three glucose level

df_glucose %>%
ggplot(mapping = aes(x = glucoseGroup, y = n, fill = factor(stroke))) +
  geom_bar(stat = "identity", position = "fill") +
  labs(title = "Percentage of subjects with Different Glucose Level", 
       x = "glucoseGroup", 
       y = "Percentage") +
  scale_fill_brewer(palette = "Paired")

```

```{r results='markup'}

table(cleanData$stroke)
#use SMOTE to create new dataset that is more balanced
cleanData <- SMOTE(stroke ~ ., cleanData, perc.over = 2000, perc.under = 200)
#minority => 2000/100 =20 so 20*(strok1Observations=416) #upsampling 
#200/100 =2, make the majority observation 2 time the minority observations #downsampling 

#view distribution of response variable in new dataset
table(cleanData$stroke)
```


```{r, results='markup'}

ggplot(data = cleanData, mapping = aes(x = stroke, y = bmi, fill=stroke)) +
  geom_boxplot()
```
With the boxplot above, subjects who have stroke has slightly lower BMI than subjects who did not have stroke. 

```{r, results='markup'}
ggplot(data = cleanData, mapping = aes(x = stroke, y = bmi, fill = stroke, color = gender)) +
  geom_boxplot() +
  labs(title = "Boxplot of BMI", 
       x = "Stroke", 
       y = "BMI") +
  scale_fill_brewer(palette = "Paired")
```
From the box plots above, there is no strong evidence of subjects having stroke with respect to BMI and gender.


**Average Glucose Level**
As seen from below graph, even though before upsampling data the higher value of glucose level were contributing to the chances of stroke i.e stroke = 1.

Average Glucose Level after upsampling

```{r results='markup'}
df_glucose <- cleanData %>%
  group_by(glucoseGroup, stroke) %>%
  count(stroke)
#grouping data into three glucose level

df_glucose %>%
ggplot(mapping = aes(x = glucoseGroup, y = n, fill = factor(stroke))) +
  geom_bar(stat = "identity", position = "fill") +
  labs(title = "Percentage of subjects with Different Glucose Level", 
       x = "glucoseGroup", 
       y = "Percentage") +
  scale_fill_brewer(palette = "Paired")
```


After Upsampling the relationship is lost. 
```{r results='markup'}
get_box_stats <- function(y, upper_limit = max(brainData$avg_glucose_level) * 1.15) {
  return(data.frame(
    y = 0.95 * upper_limit,
    label = paste(
      "Count =", length(y), "\n",
      "Mean =", round(mean(y), 2), "\n",
      "Median =", round(median(y), 2), "\n"
    )
  ))
}

ggplot(cleanData, aes(x = stroke, y = avg_glucose_level, fill = stroke)) +
  geom_boxplot() +
  labs(title = "avg glucose level distribution before sampling")+
  scale_fill_manual(values = c("#0099f8", "#e74c3c", "#2ecc71")) +
  stat_summary(fun.data = get_box_stats, geom = "text", hjust = 0.5, vjust = 0.9) +
  theme_classic()

```


```{r results='markup'}

# percent of people with and without stroke for glucose level in three categories 
ExpCatViz(
  cleanData %>% 
    select(stroke, glucoseGroup), 
  target="stroke",col=c("Blue", "red"))
```

As seen above, average glucose level or diabetic level of sugar is impacting the percentage of people with strokes and not.
The difference is huge for Diabetic Category


3)do plots
4)show missing values for bmi and smoking_status
5)filling NA for bmi
6)removie for smoking_status 
7)remove outliers
8)smote
9)plot them again : qqplot => normality  
10) Comparison of age,bmi, avg_glucose_level to stroke
11) correlation plot for numeric variables

12) Plot correlation matrix for categorical variables 
13) analysis of residence_type and drop it 
14)Heart Disease: variable + age 
15)Hypertension : variable +age
16)Ever Married
17)Smoking Status
18)Gender
19)WorkType
20)All the testing anova, chi-square
21)Conclusion : major factor, what all other factors it depends on, (logistic regression test)




