---
title: "finalEDA"
author: "Team 6"
date: "2022-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing all the relevant packages and libraries 

```{r results='markup'}
#install packages
#remotes::install_github("cran/DMwR")


#import libraries all
#library(DMwR)
``` 

Importing our dataset and checking summary
```{r results='markup'}
#importing brain dataset
data <- data.frame(read.csv("dataset.csv"))


#checking the structure of our dataset 
str(data)

#since ID won't be useful during prediction so removing that 
brainData <- select(data, -c('id'))

#checking top5 entries in our dataset 
head(brainData)
```

Checking the summary of our dataset
```{r results='markup'}
summary(brainData)
```

Since most of the varibales our categorical, converting them to factors
```{r results='markup'}
#Converting gender,hypertension,heart_disease,ever_married,work_type,Residence_type,smoking_status,stroke to factors
brainData$gender = factor(brainData$gender)
brainData$hypertension = factor(brainData$hypertension)
brainData$heart_disease = factor(brainData$heart_disease)
brainData$ever_married = factor(brainData$ever_married)
brainData$work_type = factor(brainData$work_type)
brainData$Residence_type = factor(brainData$Residence_type)
brainData$smoking_status = factor(brainData$smoking_status)
brainData$stroke = factor(brainData$stroke)
str(brainData)
summary(brainData)
```


*We can see missing values in bmi, smoking_status \
**Class Imbalance problem** \
we have 42617 observations for stroke = 0 and 783 observations for stroke =1 \
\ 
Cleaning the dataset


Removing missing Values for smoking status 
```{r results='markup'}
cleanData = brainData
cleanData[cleanData == ''] <- NA
nrow(cleanData)
cleanData <- subset(cleanData, !is.na(smoking_status))
nrow(cleanData)
```

Replacing Missing Values : BMI 
```{r results='markup'}

#removing missing values for BMI
cleanData$bmi[is.na(cleanData$bmi)] <- median(cleanData$bmi, na.rm = T) #replacing NA values with median of the remaining bmi values.
summary(cleanData)
```
Removing Outliers for BMI 
```{r results='markup'}

dim(cleanData)
quartiles_bmi <- quantile(cleanData$bmi, probs=c(.25, .75), na.rm = FALSE)
IQR_bmi <- IQR(cleanData$bmi)

 
Lower_bmi <- quartiles_bmi[1] - 1.5*IQR_bmi
Upper_bmi <- quartiles_bmi[2] + 1.5*IQR_bmi

#since all the missing values are for stroke =0 entries, we will directly drop it

cleanData <- subset(cleanData, cleanData$bmi > Lower_bmi & cleanData$bmi < Upper_bmi)

dim(cleanData)
```



Removing Outliers for Average Glucose Level
```{r results='markup'}

dim(cleanData)
quartiles_glucose <- quantile(cleanData$avg_glucose_level, probs=c(.25, .75), na.rm = FALSE)
IQR_glucose <- IQR(cleanData$avg_glucose_level)

 
Lower_glucose <- quartiles[1] - 1.5*IQR
Upper_glucose <- quartiles[2] + 1.5*IQR

#since all the missing values are for stroke =0 entries, we will directly drop it

cleanData <- subset(cleanData, cleanData$avg_glucose_level > Lower_glucose & cleanData$avg_glucose_level < Upper_glucose)

dim(cleanData)

```

Let's check the distribution for numeric variables, bmi,age, avg_glucose_level : boxplot(outlier) and histogram(normal) => outliers **while doing avg_glucose_level and bmi => groupwise analysis as well**
3)do plots

```{r results='markup'}
```

4)show missing values for bmi and smoking_status
5)filling NA for bmi
6)removie for smoking_status 
7)remove outliers


8)smote
```{r results='markup'}

table(cleanData$stroke)
#use SMOTE to create new dataset that is more balanced
cleanData <- SMOTE(stroke ~ ., cleanData, perc.over = 2000, perc.under = 200)
#minority => 2000/100 =20 so 20*(strok1Observations=416) #upsampling 
#200/100 =2, make the majority observation 2 time the minority observations #downsampling 

#view distribution of response variable in new dataset
table(cleanData$stroke)

```
9)plot them again : qqplot => normality  

```{r results='markup'}

```

10) Comparison of age,bmi, avg_glucose_level to stroke
11) correlation plot for numeric variables

12) Plot correlation matrix for categorical variables 
13) analysis of residence_type and drop it 
14)Heart Disease: variable + age 
15)Hypertension : variable +age
16)Ever Married
17)Smoking Status
18)Gender
19)WorkType
20)All the testing anova, chi-square
21)Conclusion : major factor, what all other factors it depends on, (logistic regression test)




