---
title: "initialEDA"
output: html_document
date: "2022-10-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
data <- data.frame(read.csv("dataset.csv"))

# Extracting the variables we need from the data set
brainData <- select(data, -c('id'))
head(brainData)
str(brainData)
```

Converting categorical variables as factor
```{r}
#gender,hypertension,heart_disease,ever_married,work_type,Residence_type,smoking_status,stroke
brainData$gender = factor(brainData$gender)
brainData$hypertension = factor(brainData$hypertension)
brainData$heart_disease = factor(brainData$heart_disease)
brainData$ever_married = factor(brainData$ever_married)
brainData$work_type = factor(brainData$work_type)
brainData$Residence_type = factor(brainData$Residence_type)
brainData$smoking_status = factor(brainData$smoking_status)
brainData$stroke = factor(brainData$stroke)
str(brainData)
```
```{r results='markup'}

```

```{r results='markup'}
brainData$bmi[is.na(brainData$bmi)] <- median(brainData$bmi, na.rm = T) #replacing NA values with median of the remaining bmi values.
corr_brain <- select_if(brainData, is.numeric)
head(corr_brain)
library("corrplot")
#cor_mat <- cor(corr_brain)
#corrplot(cor_mat,method="number")
#pairs(brainData)
```

```{r results='markup'}
library(ggplot2)
ggplot(brainData, aes(y=avg_glucose_level, x=stroke)) + 
  geom_boxplot() + 
  geom_boxplot( colour="orange", fill="#7777cc", outlier.colour="red", outlier.shape=4, outlier.size=2) +
  labs(title="outlier in avg glucose level  not due to stroke`", x="", y = "avg glucose level")

ggplot(brainData, aes(y=bmi, x=stroke)) + 
  geom_boxplot() + 
  geom_boxplot( colour="orange", fill="#7777cc", outlier.colour="red", outlier.shape=4, outlier.size=2) +
  labs(title="outlier in bmi  not due to stroke`", x="", y = "bmi")


```
**avg glucose level of people with stroke is higher than the max of people without stroke**

```{r results='markup'}
ggplot(data=brainData, aes(bmi)) + 
  geom_histogram(breaks=seq(66, 86, by = 1), 
                 col="red", 
                 fill="blue", 
                 alpha = .7) + # opacity
  labs(x="bmi", y="Frequency") +
  labs(title="bmi histogram using `ggplot`") 


ggplot(data=brainData, aes(age)) + 
  geom_histogram(breaks=seq(66, 86, by = 1), 
                 col="red", 
                 fill="blue", 
                 alpha = .7) + # opacity
  labs(x="age", y="Frequency") +
  labs(title="bmi age using `ggplot`") 



```


```{r results='markup'}
#bmi
ggplot(data = brainData, mapping = aes(x = work_type, y = bmi)) +
  geom_boxplot()


ggplot(data = brainData, mapping = aes(x = Residence_type, y = bmi)) +
  geom_boxplot()

ggplot(data = brainData, mapping = aes(x = ever_married, y = bmi)) +
  geom_boxplot()

ggplot(data = brainData, mapping = aes(x = smoking_status, y = bmi)) +
  geom_boxplot()

ggplot(data = brainData, mapping = aes(x = hypertension, y = bmi)) +
  geom_boxplot()


```

```{r results='markup'}
#avg_glucose_level
ggplot(data = brainData, mapping = aes(x = work_type, y = avg_glucose_level)) +
  geom_boxplot()


ggplot(data = brainData, mapping = aes(x = Residence_type, y = avg_glucose_level)) +
  geom_boxplot()

ggplot(data = brainData, mapping = aes(x = ever_married, y = avg_glucose_level)) +
  geom_boxplot()

ggplot(data = brainData, mapping = aes(x = smoking_status, y = avg_glucose_level)) +
  geom_boxplot()
#hypertension high => high avg bloood sugar
ggplot(data = brainData, mapping = aes(x = hypertension, y = avg_glucose_level)) +
  geom_boxplot()

```

```{r, results='markup'}
qqnorm(brainData$age, main="Q-Q plot of age", col="blue", ylab="age quantiles")
qqline(brainData$age)

qqnorm(brainData$bmi, main="Q-Q plot of bmi", col="blue", ylab="bmi quantiles")
qqline(brainData$bmi)

qqnorm(brainData$avg_glucose_level, main="Q-Q plot of avg_glucose_level", col="blue", ylab="avg_glucose_level quantiles")
qqline(brainData$avg_glucose_level)

```
```{r, results='markup'}
summary(brainData)
```
Number of NA values in bmi: 1462\
Number of NA values in smoking_status: 13292\
```{r, results='markup'}
library(ggplot2)
ggplot(data=brainData, aes(bmi)) +
geom_histogram(col="black",
fill="blue") +
labs(x="bmi", y="frequency") +
labs(title="plot of bmi")
```
```{r, results='markup'}
tempdf <- subset(brainData, !is.na(bmi))  # removing NA values from bmi and storing it in tempdf
density_bmi <- density(tempdf$bmi)
plot(density_bmi)

ggplot(data = tempdf, mapping = aes(y = bmi)) +
  geom_boxplot() + geom_boxplot( colour="black", fill="#7777cc", outlier.colour="red", outlier.shape=8, outlier.size=2)

```
From the boxplot, we can see that there are multiple outliers.
```{r, results='markup'}
#install.packages("moments")
library(moments)

skewness(tempdf$bmi)
```
From the density curve and skewness value, BMI is rightly skewed. So mean is not the right central tendency to fill NA values in BMI.
We will NA values in BMI using median of the data.
Note: Negative skewness value indicates left skewed and positive skewness value indicates right skewed.
The rule of thumb seems to be:

If the skewness is between -0.5 and 0.5, the data are fairly symmetrical
If the skewness is between -1 and – 0.5 or between 0.5 and 1, the data are moderately skewed
If the skewness is less than -1 or greater than 1, the data are highly skewed



```{r, results='markup'}
brainData$bmi[is.na(brainData$bmi)] <- median(brainData$bmi, na.rm = T) #replacing NA values with median of the remaining bmi values.
summary(brainData)
```

correlation plot for Numeric Variables :

```{r results='markup'}
corr_brain <- select_if(brainData, is.numeric)
head(corr_brain)
library("corrplot")
cor_mat <- cor(corr_brain)
corrplot(cor_mat,method="number")

```

Correlation plot for Factor variables
```{r results='markup'}
library(tidyverse)
df_character <- select_if(brainData, is.factor)
head(df_character)
#plot_correlation(df_character)
#M = cor(corr_brain)
#corrplot(M, method = 'number') # colorful number
#plot_c
```





**Data Analysis for average glucose level**
```{r results='markup'}
get_box_stats <- function(y, upper_limit = max(brainData$avg_glucose_level) * 1.15) {
  return(data.frame(
    y = 0.95 * upper_limit,
    label = paste(
      "Count =", length(y), "\n",
      "Mean =", round(mean(y), 2), "\n",
      "Median =", round(median(y), 2), "\n"
    )
  ))
}


ggplot(brainData, aes(x = stroke, y = avg_glucose_level, fill = stroke)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#0099f8", "#e74c3c", "#2ecc71")) +
  stat_summary(fun.data = get_box_stats, geom = "text", hjust = 0.5, vjust = 0.9) +
  theme_classic()

```
```{r results='markup'}
get_box_stats_smoking <- function(y, upper_limit = max(brainData$avg_glucose_level) * 1.15) {
  return(data.frame(
    y = 0.95 * upper_limit,
    label = paste(
      "Count =", length(y), "\n",
      "Mean =", round(mean(y), 2), "\n",
      "Median =", round(median(y), 2), "\n"
    )
  ))
}


ggplot(brainData, aes(x = smoking_status, y = avg_glucose_level, fill = stroke)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#0099f8", "#e74c3c", "#2ecc71","#2ecc71")) +
  stat_summary(fun.data = get_box_stats_smoking, geom = "text", hjust = 0.5, vjust = 0.9) +
  theme_classic()


# Basic Scatterplot Matrix
pairs(~avg_glucose_level+bmi+age,data=brainData,
   main="Simple Scatterplot Matrix")

```
```{r results='markup'}
# plot the distribution of salaries 
# by rank using kernel density plots
ggplot(brainData, 
       aes(x = avg_glucose_level, 
           fill = stroke)) +
  geom_density(alpha = 0.4) +
  labs(title = "avg glucose level distribution by rank")

ggplot(brainData, 
       aes(x = avg_glucose_level, 
           fill = smoking_status)) +
  geom_density(alpha = 0.4) +
  labs(title = "avg glucose level distribution by rank")
ggplot(brainData, 
       aes(x = avg_glucose_level, 
           fill = Residence_type)) +
  geom_density(alpha = 0.4) +
  labs(title = "avg glucose level distribution by rank")
```


```{r results='markup'}
library(ggcorrplot)
model.matrix(~0+., data=brainData) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)

```


checking smoking status
Based on the correlation matrix, 
smoked is closely related to age, work_type=private
Never Smoke closely related to age, ever_married =yes,private, bmi
fSMOKE bmi, married=yes, age

```{r results='markup'}
smoked_private_yes_dataset <- subset(brainData, work_type=="Private" & ever_married=="Yes" )
smoked_private_no_dataset <- subset(brainData, work_type=="Private" & ever_married=="No" )
ggplot(smoked_private_yes_dataset, 
       aes(x = age, 
           fill =smoking_status )) +
  geom_density(alpha = 0.4) +
  labs(title = "age and smoking status evrmarried=yes ")
ggplot(smoked_private_no_dataset, 
       aes(x = age, 
           fill =smoking_status )) +
  geom_density(alpha = 0.4) +
  labs(title = "age and smoking status evrmarried=no ") #age <25 => formerly smokes 
#30-40 : smokes
#40-80 : formerly smokes
#could be marked as formerly smokes
```

```{r results='markup'}
library(ggplot2)
#data(Salaries, package="carData")

# plot experience vs. salary 
# (color represents rank and size represents service)
ggplot(brainData, 
       aes(x = bmi, 
           y = age, 
           size = ever_married, 
           color = smoking_status)) +
  geom_point(alpha = .6) +
  labs(title = "Academic salary by rank, years of service, and years since degree")
```


```{r results='markup'}
missing_smoke <- subset(brainData, smoking_status=="" & age < 10)
less_30 <- subset(brainData, age < 30)
missing_smoke
ggplot(brainData, 
       aes(x = age, 
           fill = smoking_status)) +
  geom_density(alpha = 0.4) +
  labs(title = "avg glucose level distribution by rank")

ggplot(less_30, 
       aes(x = age, 
           fill = smoking_status)) +
  geom_density(alpha = 0.4) +
  labs(title = "avg glucose level distribution by rank")
```


```{r results='markup'}
get_box_stats_ <- function(y, upper_limit = max(brainData$age) * 1.15) {
  return(data.frame(
    y = 0.95 * upper_limit,
    label = paste(
      "Count =", length(y), "\n",
      "Mean =", round(mean(y), 2), "\n",
      "Median =", round(median(y), 2), "\n"
    )
  ))
}


ggplot(brainData, aes(x = smoking_status, y = age, fill = work_type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("orange", "black", "blue","green","yellow")) +
  stat_summary(fun.data = get_box_stats_, geom = "text", hjust = 0.5, vjust = 0.9) +
  theme_classic()
ggplot(brainData, aes(x = smoking_status, y = age, fill = ever_married)) +
  geom_boxplot() +
  scale_fill_manual(values = c("orange", "black", "blue","green","yellow")) +
  #stat_summary(fun.data = get_box_stats_, geom = "text", hjust = 0.5, vjust = 0.9) +
  theme_classic()


ggplot(brainData, aes(x = smoking_status, y = avg_glucose_level, fill = work_type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("orange", "black", "blue","green","yellow")) +
  #stat_summary(fun.data = get_box_stats_, geom = "text", hjust = 0.5, vjust = 0.9) +
  theme_classic()
ggplot(brainData, aes(x = smoking_status, y = avg_glucose_level, fill = ever_married)) +
  geom_boxplot() +
  scale_fill_manual(values = c("orange", "black", "blue","green","yellow")) +
  #stat_summary(fun.data = get_box_stats_, geom = "text", hjust = 0.5, vjust = 0.9) +
  theme_classic()
```


**Data Analysis for Smoking Status**

```{r results='markup'}
brainData[brainData == ''] <- NA
no_na_for_smoking_status <- subset(brainData, !is.na(smoking_status))
no_na_for_smoking_status
ggplot(no_na_for_smoking_status, aes(smoking_status, age, color = stroke)) + 
  geom_boxplot() + 
  ggtitle("Distribution for age vs smoking_status on stroke")
select_smoke <- select(no_na_for_smoking_status, c('smoking_status','stroke'))
smoke_effect <- select_smoke %>% group_by(smoking_status, stroke) %>% count(stroke)
ggplot(smoke_effect, aes(smoking_status, n, fill = stroke)) + 
  geom_bar(stat = 'identity', position='fill') + 
  ggtitle("Distribution for smoking_status on stroke")

```
After the analysis of smoking status against stroke where `0` is `No` and `1` is `Yes`, we notice that smoking status does not really have an effect on stroke. According to the graph, we see that people who formerly smoked have more tendency to get strokes than people who never smoked and are still smoking. The boxplot distribution then shows that age is indeed has a great effect on stroke, we see that people approximately above 45 and formerly smoked have strokes under a normal distribution and younger people across all distribution do not have strokes.


```{r, results='markup'}
#install.packages("DataExplorer")
library(DataExplorer)
#checking the unknown value in smoking_status
plot_bar(brainData$smoking_status)

#age continuous to discrete 
#0-14 
#15-44
#45-64
# >64
brainData$ageGroup=cut(brainData$age, c(0, 14, 44, 64, Inf), c("0-14", "15-44", "45-64", ">64"), include.lowest=TRUE)
#brainData$ageGroup

#plot_bar(brainData$ageGroup,by =brainData$smoking_status)
#install.packages("ggstatsplot")
library(ggstatsplot)
ggbarstats(data=brainData,x=ageGroup,y=smoking_status,label="both")
#maximum are in category of 0-14 so will fill the entries with never smokes 

```


```{r results='markup'}
brainData$glucoseGroup=cut(brainData$avg_glucose_level, c(0, 114, 140, Inf), c("Normal", "Prediabetes", "Diabetes"), include.lowest=TRUE)
#0-114 : Normal glucose level
#114-140 : Prediabetes
#140- infinity : Diabetes 
ggbarstats(data=brainData,x=glucoseGroup,y=ageGroup,label="both")
ggbarstats(data=brainData,x=glucoseGroup,y=smoking_status,label="both")


#glucoseGroup to stroke relations
ggbarstats(data=brainData,y=glucoseGroup,x=stroke,label="both")
#people with diabetes have a high chance of getting stroke 

```


```{r results='markup'}
brainData$bmiGroup=cut(brainData$bmi, c(0, 18.5, 24.9, 29.9,Inf), c("Underweight", "Healthy", "Overweight","Obese"), include.lowest=TRUE)

#If your BMI is less than 18.5, it falls within the underweight range.
#If your BMI is 18.5 to 24.9, it falls within the normal or Healthy Weight range.
#If your BMI is 25.0 to 29.9, it falls within the overweight range.
#If your BMI is 30.0 or higher, it falls within the obese range.

str(brainData)

ggbarstats(data=brainData,y=bmiGroup,x=stroke,label="both")
```
Summary Analysis of Average Glucose level \



Histogram and Box Plot Distribution of Average Glucose Level

```{r results='markup'}
#histogram for Average Glucose Level
library(ggplot2)
ggplot(data=brainData, aes(avg_glucose_level)) +
geom_histogram(col="black",
fill="blue") +
labs(x="Avg Glucose Level", y="frequency") +
labs(title="Distribution for Average Glucose Level")

#Boxplot for average glucose level
ggplot(data = brainData, mapping = aes( y = avg_glucose_level)) +
  geom_boxplot()

```
```{r results='markup'}
#install.packages("flextable")
library(flextable)
library(SmartEDA) 
ExpNumStat(brainData, by="GA", gp="stroke", Outlier=TRUE, Qnt = c(.25, .75), round = 2) %>% flextable()

#outliers
#library(performance)
#plot(check_outliers(brainData$avg_glucose_level, method = "zscore"))
```
```{r results='markup'}

#removing outliers for Glucose from the data 
#brainData_removedOutlierGlucose = 

quartiles <- quantile(brainData$avg_glucose_level, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(brainData$avg_glucose_level)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
#since all the missing values are for stroke =0 entries, we will directly drop it

data_no_outlier = brainData
#data_no_outlier$avg_glucose_level <- replace(data_no_outlier$avg_glucose_level,data_no_outlier$avg_glucose_level > Upper, NA)

 
data_no_outlier <- subset(brainData, brainData$avg_glucose_level > Lower & brainData$avg_glucose_level < Upper)
dim(brainData)
dim(data_no_outlier)



#Histogram for average glucose level after removing outlier
ggplot(data=data_no_outlier, aes(avg_glucose_level)) +
geom_histogram(col="black",
fill="blue") +
labs(x="Avg Glucose Level", y="frequency") +
labs(title="Distribution for Average Glucose Level")

#Boxplot for average glucose level after removing outlier
ggplot(data =data_no_outlier, mapping = aes( y = avg_glucose_level)) +
  geom_boxplot()
ExpNumStat(data_no_outlier, by="GA", gp="stroke", Outlier=TRUE, Qnt = c(.25, .75), round = 2) %>% flextable()
```

Checking Normality using qq plot

```{r results='markup'}
#qq plot for checking normality 
plot_qq(brainData, by = "stroke") 
plot_qq(data_no_outlier, by = "stroke")

#Shapiro-Wilk Normality test.
#install.packages("ggpubr")
#library(ggpubr)
#ggqqplot(brainData, "avg_glucose_level", facet.by = "stroke")
ggqqplot(brainData, "avg_glucose_level")
ggqqplot(data_no_outlier, "avg_glucose_level")
```

As seen from above plots, average glucose level curve is more close to a normal distribution after removing outliers.

```{r results='markup'}

# Fit the model
model <- glm( stroke ~ avg_glucose_level , data = brainData, family = binomial)
# Summarize the model
summary(model)
#(model.null - model.proposed)/model.null

# Fit the model
model1 <- glm( stroke ~ glucoseGroup , data = brainData, family = binomial)
# Summarize the model
summary(model1)

(model1.null - model1.proposed)/model1.null
```
Equation would be : 
p = exp(-5.15 + 0.010*`avg_glucose_level`)/ [1 + exp(-5.15 + 0.010*`avg_glucose_level`)]
With the low value of p we can reject our null hypothesis and thus, avg_glucose_level is closely related to stroke

```{r results='markup'}
# Fit the model
model2 <- glm( stroke ~ glucoseGroup , data = brainData, family = binomial)
# Summarize the model
summary(model2)

model2.null <- model2$null.deviance/-2 
model2.proposed <- model2$deviance/-2
(model2.null - model2.proposed)/model2.null

```
log(odds) of stroke that a normal person has stroke are : -4.28507 +  0.09157x0 + 1.20138x0 
log(odds) of stroke that a prediabetes person has stroke are : -4.28507 +  0.09157x1 + 1.20138x0 
log(odds) of stroke that a diabetes person has stroke are : -4.28507 +  0.09157x0 + 1.20138x1


Checking for all variables
```{r results='markup'}
# Fit the model
model2 <- glm( stroke ~. , data = brainData, family = binomial)
# Summarize the model
summary(model2)

model2.null <- model2$null.deviance/-2 
model2.proposed <- model2$deviance/-2
(model2.null - model2.proposed)/model2.null
```




```{r results='markup'}
#concise box plot for numeric variables
plot_boxplot(brainData, by = "stroke")
plot_boxplot(data_no_outlier, by = "stroke")
```
Converting Glucose into Categorical variable and testing the results 
```{r results='markup'}
library(SmartEDA) 
# percent of people with and without stroke for glucose level in three categories 
ExpCatViz(
  brainData %>% 
    select(stroke, glucoseGroup), 
  target="stroke",col=c("Blue", "red"))
#density plot of 3 numeric groupp age,avg_glucose_level, bmi
#plot_density(brainData)

#install.packages("flextable")
#checking correlation
ggcorrmat(
  data   = brainData,
  type   = "np",
  output = "dataframe"
) %>% 
  mutate_if(is.numeric, ~round(., 3)) %>% 
  flextable()
```

Checking Plots for Correlation
```{r results='markup'}
#install.packages("dlookr")
#library(dlookr)
#checking various plots for correlation
#plot_correlate(brainData, method = "kendall")
#ggscatterstats(
#  data = brainData,
#  x = age,
#  y = avg_glucose_level,
#  type = "np" # try the "robust" correlation too! It might be even better here
  #, marginal.type = "boxplot"
#)

#library(PerformanceAnalytics)
chart.Correlation(brainData %>% select(-glucoseGroup), method = "kendall") 

```


Removing Outlier 
```{r results='markup'}
#brainData[brainData == ''] <- NA
#no_na_for_smoking_status <- subset(brainData, !is.na(smoking_status))
#no_na_for_smoking_status
#
#without avg_glucose_level outliers : data_no_outlier


quartiles <- quantile(brainData$avg_glucose_level, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(brainData$avg_glucose_level)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
#since all the missing values are for stroke =0 entries, we will directly drop it

data_no_outlier = brainData
data_no_outlier <- subset(brainData, brainData$avg_glucose_level > Lower & brainData$avg_glucose_level < Upper)
dim(brainData)
dim(data_no_outlier)



#data_no_outlier[data_no_outlier == ''] <- NA
#data_no_outlier<- subset(data_no_outlier, !is.na(smoking_status))
data_no_outlier <- subset(data_no_outlier, data_no_outlier$smoking_status != '' )
dim(data_no_outlier)
```
```{r results='markup'}
quartiles_bmi <- quantile(data_no_outlier$bmi, probs=c(.25, .75), na.rm = FALSE)
IQR_bmi <- IQR(data_no_outlier$bmi)
Lower_bmi <- quartiles_bmi[1] - 1.5*IQR_bmi
Upper_bmi <- quartiles_bmi[2] + 1.5*IQR_bmi
#since all the missing values are for stroke =0 entries, we will directly drop it
data_no_outlier <- subset(data_no_outlier, data_no_outlier$bmi > Lower_bmi & data_no_outlier$bmi< Upper_bmi)
#dim(brainData)
dim(data_no_outlier)
#
yes_outlier <- subset(data_no_outlier, data_no_outlier$stroke == 1)
no_outlier <- subset(data_no_outlier, data_no_outlier$stroke == 0)
nrow(yes_outlier)
nrow(no_outlier)
```



```{r, results='markup'}
df_hypertension_stroke <- data1 %>%
  group_by(hypertension, stroke) %>%
  count(stroke)
#xkabledply(df_hypertension_stroke)
df_hypertension_stroke
```

```{r, results='markup'}
df_hypertension_stroke %>%
ggplot(mapping = aes(x = hypertension, y = n, fill = factor(stroke))) +
  geom_bar(stat = "identity", position = "fill") +
  labs(title = "Percentage of subjects with Hypertension vs. without Hypertension", 
       x = "Hypertension", 
       y = "Percentage") +
  scale_fill_brewer(palette = "Paired")
```

```{r, results='markup'}
str(brainData)

df_count_hyptension_with_age = group_by(brainData, age, hypertension) %>% count(hypertension)
#df_count_hyptension_with_age

hypertension_positive = subset(df_count_hyptension_with_age, df_count_hyptension_with_age$hypertension==1)
#hypertension_positive
ggplot(hypertension_positive, mapping = aes(x=age, y=n)) + geom_point(alpha=0.5) + geom_smooth(method="loess",span=0.3)+
  labs(title="trend of number of hypertension cases with age", subtitle="Age distibution",x="Age",y="Number of subjects with hypertension")

```


As we can see from the graph that the number of hypertension cases increases with the age due to multiple biological factors.
From the above charts, we see that the number of strokes is correlated to the hypertension status of the subject. But hypertension is in-turn correlated to age.


**Data Analysis for Heart Disease**

```{r results='markup'}
select_heart_disease <- select(brainData, c('heart_disease','stroke'))
heart_effect <- select_heart_disease %>% group_by(heart_disease, stroke) %>% count(stroke)
ggplot(heart_effect, aes(heart_disease, n, fill = stroke)) + 
  geom_bar(stat = 'identity', position='fill') + 
  ggtitle("Distribution for heart disease on stroke")
ggplot(brainData, aes(heart_disease, age, color = stroke)) + 
  geom_boxplot() + 
  ggtitle("Distribution for age vs heart_disease on stroke")

count_heart_disease_with_age = group_by(brainData, age, heart_disease) %>% count(heart_disease)

heart_disease_positive = subset(count_heart_disease_with_age, count_heart_disease_with_age$heart_disease==1)

ggplot(heart_disease_positive, mapping = aes(x=age, y=n)) +
  geom_point() + 
  geom_smooth(method="loess",span=0.3) +
  labs(title="Trend of number of heart disease cases with age", subtitle="Age distribution",x="Age",y="Number of subjects with heart disease")
```
 
* From the plots above we see that heart disease does not really have a strong impact on stroke unlike age which has a big factor on stroke. We then see that an increase in age comes with an increase in heart disease. This means that people in the higher age bracket are prone to heart diseases than people in the lower age bracket, this is the relation between heart disease and age.