---
title: "Modelling"
author: "Team 6"
date: "2022-11-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing the required libraries
```{r}
library(caret)
library(dplyr)
library(lessR)
```


##Loading both the datasets 
```{r results='markup'}
beforeSMOTE <- data.frame(read.csv("beforeSmote.csv"))
afterSMOTE <- data.frame(read.csv("afterSmote.csv"))

```

##Factoring the Categorical variables as before 
```{r results='markup'}
#factoring the dataset
beforeSMOTE$gender = factor(beforeSMOTE$gender)
beforeSMOTE$hypertension = factor(beforeSMOTE$hypertension)
beforeSMOTE$heart_disease = factor(beforeSMOTE$heart_disease)
beforeSMOTE$ever_married = factor(beforeSMOTE$ever_married)
beforeSMOTE$work_type = factor(beforeSMOTE$work_type)
beforeSMOTE$Residence_type = factor(beforeSMOTE$Residence_type)
beforeSMOTE$smoking_status = factor(beforeSMOTE$smoking_status)
beforeSMOTE$stroke = factor(beforeSMOTE$stroke)
summary(beforeSMOTE)

#factoring the dataset
afterSMOTE$gender = factor(afterSMOTE$gender)
afterSMOTE$hypertension = factor(afterSMOTE$hypertension)
afterSMOTE$heart_disease = factor(afterSMOTE$heart_disease)
afterSMOTE$ever_married = factor(afterSMOTE$ever_married)
afterSMOTE$work_type = factor(afterSMOTE$work_type)
afterSMOTE$Residence_type = factor(afterSMOTE$Residence_type)
afterSMOTE$smoking_status = factor(afterSMOTE$smoking_status)
afterSMOTE$stroke = factor(afterSMOTE$stroke)
summary(afterSMOTE)

cols <-  hcl.colors(length(levels(beforeSMOTE$stroke)), "Fall")
PieChart(stroke, data = beforeSMOTE, hole = 0,
         fill = cols,
         labels_cex = 0.6)
```

### After SMOTE
```{r}
cols <-  hcl.colors(length(levels(afterSMOTE$stroke)), "Fall")
PieChart(stroke, data = afterSMOTE, hole = 0,
         fill = cols,
         labels_cex = 0.6)
```


## Setting the variable for checking 
```{r}
#set it to beforeSMOTE/afterSMOTE

#modellingDataset <- beforeSMOTE
modellingDataset <- afterSMOTE
```


## Splitting the dataset into train and test 
```{r}
set.seed(123)  
split1<- sample(c(rep(0, 0.7 * nrow(modellingDataset)), rep(1, 0.3 * nrow(modellingDataset))))
table(split1) 
#train split
train <- modellingDataset[split1 == 0, ]
#test split
test <- modellingDataset[split1== 1, ] 
```

###Logistic Regression
```{r results='markup'}
# fit logistic regression model 
logitModel <- glm(stroke~ ., 
                        data = train, 
                        family = binomial())

summary(logitModel)
```
Significant variables in case of beforeSMOTE 

* age 
* hypertension 
* heart_disease 
* avg_glucose_level 

Significant variables after applying SMOTE

* genderMale
* age
* hypertension1
* heart_disease1
* work_typeGovt_job
* smoking_statussmokes

#Predicting accuracy 
```{r }
logitModelPred <- predict(logitModel, test, type = "response")
# plot of probabilities
plot(logitModelPred, 
     main = "Scatterplot of Probabilities of stroke (test data)", 
     xlab = "Patients ", ylab = "Predicted Probability of stroke")
```




